---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(knitr)
library(modelsummary)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

```{r}
eggs_data <- read_parquet("./data/02-analysis_data/cleaned_data.parquet")
head(eggs_data)

```

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

## Measurement
	
Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

## Predictor variables


# Results

```{r}
#|label: fig-ttlprice-ppu  
#| include: false
#| warning: false
#| message: false
#| fig-cap: Higher Quantity Egg Products Are More Expensive
# removing outliers

eggs_data <- eggs_data |>
  filter(price_per_unit_numeric < quantile(eggs_data$price_per_unit_numeric, 0.75) + IQR(eggs_data$price_per_unit_numeric)) |>
  filter(price_per_unit_numeric > quantile(eggs_data$price_per_unit_numeric, 0.25) - IQR(eggs_data$price_per_unit_numeric))

head(eggs_data)

eggs_data |> 
  ggplot(aes(price_per_unit_numeric, current_price, color = numerical_units)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "Current Price", y = "Price Per Unit") +
  scale_color_continuous(name = "Egg Quantity")
```

From @fig-ttlprice-ppu, there appears to be a weak linear relationship between current egg product price and the price per a single egg unit. However, after including egg quanitity, the relationship between current egg product price and egg quantity as a predictor of cost of an egg per a single unit is quite strong.

```{r}
#|label: fig-brand-cost
#| include: false
#| warning: false
#| message: false
#| fig-cap: Some Egg Brands are More Expensive than Others
# removing outliers

brand_avg_quan <- eggs_data |> group_by(brand) |> summarise(avg_brand_quantity = mean(numerical_units))

eggs_data <- eggs_data |> 
  inner_join(brand_avg_quan, by = "brand")
eggs_data |> 
  ggplot(aes(brand, current_price, color = avg_brand_quantity.x)) +
  geom_boxplot(outlier.shape = NA) +
  theme_minimal() +
  labs(y = "Current Price", x = "Egg Brand") + 
  scale_color_continuous(name = "Average Egg Quantity") +
  coord_flip()
```

@fig-brand-cost shows that some brands have more expensive egg quantities than others. For example, Rabbit River Farms is more expensive than President's choice. Also, it is obvious that most of the brands offer a variety of products since the IQR of their prices are quite large.

# Discussion

## Correlation vs. Causation {#sec-correcause}

From @fig-brand-cost, some companies eggs are more important than other, for instance Gray Ridge is more expensive than Burnbrae Farms. However, this is not true since Golden Valley sells more eggs per one batch than Burnbrae Farms. While this is not a traditional cause of correlation vs. causation this is a similar siutation. Since only considering the relationship between the egg brand and the price per batch, we might think some brands are more expensive than others, but when considering the egg qunatity and price per single egg (@fig-ttlprice-ppu), we realize a different relationship olds true.

## Missing Data {#sec-missingdata}

In this analysis, the old egg price variable was removed. This is because almost all of the data entries did not include this column. This limited our analysis since it made it quite difficult to analyze how specfic egg product prices changed over time.

## Sources of Bias {#sec-bias}

In @fig-ttlprice-ppu, by only looking at the linear model lines, there does not appear to be a strong relationship between current egg price and the price per unit. However, by looking at the graph in general, there appear to be four different positive, strong, linear relationships between current egg price and price per egg. This suggests that there is another variable affecting the reltionship, which based on the graph is the quantity of eggs in that batch. Thus, most conclusions regarding current egg price must be interpreted in a manner that accounts for the number of eggs in that batch.

## Weaknesses And Next Steps

The biggest issue in this paper is the limitation of the data. The team had limited numerical data and it was only regarding egg price and egg quantity. In the future, we would gather purchase data to see how changes in egg prices could change consumer spending habits. We would also like to incorporate other historical data like interest changes, tax changes, and how those affect consumer egg spending habits and egg prices.

\newpage

\appendix

# Appendix {-}

# Additional data details

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


